\begin{abstract}\thispagestyle{plain}
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}
There is a constantly growing amount of data and business is evolving to be data driven. Data scientists, machine learners and statisticians are rare and expensive. Not every organization can afford to hire such specialists, and available standard models and software might not be suitable. Hence, we propose a data model driven architecture that enables application programmers and domain experts to easily customize standard machine learning algorithms and techniques. This approach is based on a framework that offers probabilistic inference algorithms in the form of entity-relationship models. We develop a technique to transform any given probabilistic model in plate notation into a well-formed entity-relationship model. Customization happens by integrating inference results with domain specific data and performing analysis afterwards, rather than customizing the probabilistic model. The integration is done on a data model basis and can easily be performed by anyone familiar with entity-relationship models.

We discuss how our data model driven approach stands out in terms of usability, comprehensibility and flexibility to end users. However, the long term success of our approach relies on the data science community which supplies algorithms for the framework and transforms the corresponding probabilistic models to entity-relationship models. The next step [TODO: different phrase, cuz no previous step] is to implement a prototype and conduct experiments or user studies. A final implementation could either be stand-alone or part of some existing software suite.
\end{abstract}
